Status of Global Unassessed Fisheries will Remain Highly Uncertain without Better Data

### 2021-01-19 documents/ovando-etal-assessing-global-fisheries.Rmd
seems like that did it

Costello et al. 2016 [@costello2016a] found similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be overly-optimistic, and the Northeast Pacific should be better by comparison. The superensemble method [@anderson2017a] used in Rosenberg et al. 2018 [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific. The Pauly 2007 [@pauly2007] catch-based approach finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)). The non-RAM results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status. However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions is concerning.

@@

Where $B_t$ is biomass at time *t*, *K* is carrying capacity ,*r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of *B~MSY~/K* to shift. When *m* is two *B~MSY~ / K* = 0.5, lower values of *m* shift the production function left, higher values right. $\pmb{c}$ is a vector catches, and $\pmb{p}$ is vector of process errors. Growth rates can become unrealistically large when the population reaches low sizes under the Pella-Tomlinson model. We deal with this problem by following the methods described in [@winker2018]. to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.

We allow for process error *p* (in the manner of the stochastic stock reduction analysis (SRA) suggested by Walters *et al.* 2006 [@walters2006]). This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average. Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see [@thorson2018]).

Process error *p* is assumed to be log-normally distributed, such that

@@

All of our estimates are Bayesian in nature. sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent). Unless there is an abundance index to fit to the model runs as a stock reduction analysis. A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery. sraplus allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~. We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history. Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected. The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status. Any model results that are not based on an index of abundance are estimated through stock reduction analysis. All stock reduction analysis style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.

@@

The full list of estimable parameters are listed in Table.\@ref(tab:param-tab). *r* and *K* are the only two parameters that are always estimated. Estimation of every other parameter can be turned on or off. When estimation is turned off estimable parameters are fixed at their initial values, which can either be set to model defaults or specified by the user. For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *K*, $\sigma_{obs}$, $\gamma$, and *B0*. *q* is also estimated when needed.

```{r param-tab}
cd <- data_frame(Parameter = "Carrying Capacity", Abbreviation = "$K$", `Default Prior` = "Prior predictive tuning") %>%
  rbind(c("Growth rate", "$r$", "Thorson, 2020 [@thorson2020] updated by prior predictive tuning")) %>%
  rbind(c("Shape parameter", "$m$", "Drawn from Thorson et al. 2012 (@thorson2012)")) %>%
  rbind(c("Catchability", "$q$", "$logn(1e^{-3}, 1)$")) %>%
  rbind(c("Observation Error", "$\\sigma_{obs}$", "$logn(.05,1)$")) %>%
  rbind(c("Ratio of process to observation error", "$\\gamma$", "$logn(.5,0.25)$")) %>%
  rbind(c("Initial State", "$B0$", "Posterior probability dist. of catch-based regressions"))

knitr::kable(
  cd,
  align = "l",
  escape = F,
  row.names = F,
  caption = "Name, abbreviations, and priors distribution for parameters potentially estimated by sraplus in this manuscript.",
  format = "pipe"
) %>%
  kableExtra::kable_styling(full_width = T, font_size = 9) %>%
  kableExtra::row_spec(0, bold = T)
```

@@

The shape parameter is usually not reliably estimable given available data for surplus production models, however, Thorson et al. 2012 [@thorson2012] provides estimate the ratio of *B~MSY~* to *K* for many fish taxa. While we estimate *m* by default throughout the results presented here, we use highly informative priors for the shape parameter based on Thorson et al. 2012 [@thorson2012] for the genus of the species in question.

We address two critical features of prior use in sraplus below: tuning of the prior-predictive distribution and translation of outside data into priors usable by sraplus.

@@

Suppose that the only thing we observe from a fishery is a catch history. Assuming only Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past. In the absence of any data to fit to, the a stock reduction algorithm  (SRA) works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history and model structure. This creates a problem for the Bayesian nature of our analysis. Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*. Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step. Doing so places two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context).

@@

Borel's Paradox poses a particular problem  for the SRA version of sraplus. Due to the fact that there are more ways for a fishery to be relatively unexploited than for a fishery to be close to collapse but not collapsed, in this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although other modeling choices can result in a net negative bias in stock status, @free2020). This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the the posterior distributions of stock status resulting from the SRA algorithm.

@@

Along with allowing users to supply their own priors, the sraplus package contains three built-in methods for converting information on stock status from additional outside data into a form usable as a stock status prior by sraplus. We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database. We then trained a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types. Given values of these variables for a new fishery, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus for new fisheries. For example, given data on SAR or FMI scores, together with a catch history, sraplus uses these regressions to convert those SAR and FMI values into priors on B/B~MSY~ or F/F~MSY~ in the most recent year of the fishery usable by sraplus (See Supplementary Information).

All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017]. The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus. Given the scope of this analysis, we do not claim that the regressions used here are the best possible model relating these data with the fishery status indicators of interest. Rather, each regression was tested to ensure that it is unlikely that, given the same data, an alternative model would perform substantially better than those presented here.

@@

We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status. From there, we assess the performance of models fit using different kinds of broadly available data: combinations of catches, effort, Fisheries Management Index scores, and swept area ratio (SAR) values (Table.\@ref(tab:data)).

@@

As a case study, we selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores. We then paired effort data at the resolution of year, country, and FAO statistical area from Rousseau et al. 2019 [@rousseau2019] to each stock. As a benchmark, we first estimated stock status for these case study fisheries using the CMSY [@froese2017] method, as this has become one of the most widely used catch-only models currently available. We then used stock-specific data on SAR and FMI to generate priors on F/F~MSY~ for each of the stocks, which were then passed to sraplus. Lastly, we used the reconstructed effort data [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values. While CMSY systemically overestimated fishing mortality rates and underestimated stock status, use of the SAR, FMI, and effort data produced substantially more accurate results for both B/B~MSY~ and F/F~MSY~ (Fig.\@ref(fig:cs-plot)).

>>>>

<!-- Costello et al. 2016 [@costello2016a] found similar rankings of regions in terms of stock status as RAM, but their estimate of state of fisheries in the Mediterranean/Black Sea regions and Southeast Asia seem to be overly-optimistic, and the Northeast Pacific should be better by comparison. The superensemble method [@anderson2017a] used in Rosenberg et al. 2018 [@rosenberg2018] demonstrates the same problem, with stocks in Southeast Asia estimated as doing better than the Northeast Atlantic or Northeast Pacific. The Pauly 2007 [@pauly2007] catch-based approach finds the stocks of Southeast Asia in much better condition than those in the Northeast Pacific or Northeast Atlantic (Table.\@ref(tab:status-tab)). The non-RAM results in Table.\@ref(tab:status-tab) include both formally assessed and unassessed stocks, and as such we would expect them to differ broadly in their estimates of regional stock status, particularly since we might expect unassessed stocks to be less rigorously managed and by extension have poorer stock status. However, the lack of consistency across heavily assessed regions such as the Northeast Pacific, and the lack of contrast in stock status between heavily and lightly managed regions is concerning. -->

@@

Where $B_t$ is biomass at time *t*, *K* is carrying capacity ,*r* is the intrinsic growth rate, *m* is the scaling parameter that allows for the ratio of *B~MSY~/K* to shift. When *m* is two *B~MSY~ / K* = 0.5, lower values of *m* shift the production function left, higher values right.  The shape parameter *m* is usually not reliably estimable given available data for surplus production models, however, Thorson et al. 2012 [@thorson2012] provides estimate the ratio of *B~MSY~* to *K* for many fish taxa. We fix the shape parameter based on the distributions reported in Thorson et al. 2012 [@thorson2012] for the genus of the species in question.

 $\pmb{c}$ is a vector catches, and $\pmb{p}$ is vector of process errors. Growth rates can become unrealistically large when the population reaches low sizes under the Pella-Tomlinson model. We deal with this problem by following the methods described in [@winker2018]. to reduce the production of the population when it falls below a threshold of 25% of carrying capacity.

We allow for process error *p* (in the manner of the stochastic stock reduction analysis (SRA) suggested by Walters *et al.* 2006 [@walters2006]). Process error *p* is assumed to be log-normally distributed, such that

<!-- This allows the population dynamics to deviate from the exact values given by the Pella-Tomlinson operating model, while still conforming to the assumptions of this model on average. Incorporation of process errors is useful for two reasons: (1) when you have an abundance index, process errors can reduce bias arising from lack of fit in a deterministic SRA whenever dynamics are poorly explained by catch-history alone, and (2) with or without an abundance index (or other info), the stochastic portion is necessary to get good uncertainty intervals (i.e., with close to nominal coverage, see [@thorson2018]). -->

@@

All of our estimates are Bayesian in nature. sraplus can be run in two forms: either as a stock reduction analysis [SRA, @walters2006], or fit to an index of abundance (fishery dependent or independent) using Hamiltonian Monte Carlo []. Unless there is an abundance index to fit to the model runs as a stock reduction analysis. A stock reduction analysis works by specifying prior distributions on population parameters and critically the recent state of the fishery. sraplus allows users to specify the most recent status in units of depletion, B/B~MSY~, F, or F/F~MSY~. We then sample from the prior distributions of the population model parameters and apply those to the production model, along with the catch history. Any run that results in the collapse of the population (catch greater than biomass in any time step) is immediately rejected. The remaining viable draws from the prior distributions are sampled in proportion to the supplied prior on recent stock status. Any model results that are not based on an index of abundance are estimated through stock reduction analysis. All stock reduction analysis style runs in our paper sampled 2,000 draws of the prior-predictive distribution from a total of 1e6 candidate draws.  For our main sets of results (everything excluding the value of information analysis), the estimated parameters are *r*, *K*, $\sigma_{obs}$, $\gamma$, and *B0*. *q* is also estimated when needed.

@@

<!-- The full list of estimable parameters are listed in Table.\@ref(tab:param-tab).  -->

<!-- *r* and *K* are the only two parameters that are always estimated. Estimation of every other parameter can be turned on or off. When estimation is turned off estimable parameters are fixed at their initial values, which can either be set to model defaults or specified by the user. -->

@@

<!-- We address two critical features of prior use in sraplus below: tuning of the prior-predictive distribution and translation of outside data into priors usable by sraplus. -->

@@

<!-- Suppose that the only thing we observe from a fishery is a catch history. Assuming only Pella-Tomlinson population dynamics, the only thing we can learn from this catch history alone is the set of model parameters that ensure that the population still exists and never collapsed in the past.  -->

In the absence of any data to fit to, the a stock reduction algorithm  (SRA) works by assuming that we know current stock status, and then finds feasible parameters to satisfy that belief given a catch history and model structure. This creates a problem for the Bayesian nature of our analysis. Consider a production model with two parameters, a growth rate *r* and a carrying capacity *K*. Once we specify prior distributions on *r* and *K*, and then apply these distributions to our model (the shape of the production function along with the catch histories), we have implicitly provided a prior on the status of the stock in all time periods, since each unique combination of *r* and *K* together with the model and the catch history produces a deterministic stock status in each time step. Doing so places two priors on recent stock status: one implicit prior through the population parameter priors, and one explicit through the users perception of recent stock status, creating a problem termed Borel's Paradox (See Poole and Raftery 2000 [@poole2000] and references therein for a discussion of Borel's Paradox in a fisheries context).

@@

<!-- Borel's Paradox poses a particular problem  for the SRA version of sraplus. Due to the fact that there are more ways for a fishery to be relatively unexploited than for a fishery to be close to collapse but not collapsed, in this context Borel's Paradox causes the posterior distribution of stock status to be positively biased relative to the supplied prior (although other modeling choices can result in a net negative bias in stock status, @free2020). This process can also make it easy for users to accidentally supply very informative priors on stock status, without realizing that choices relating to population biology priors that may appear independent of stock status are in fact dictating the the posterior distributions of stock status resulting from the SRA algorithm. -->

@@

Along with allowing users to supply their own priors, the sraplus package contains three built-in methods for converting information on stock status from additional outside data into a form usable as a stock status prior by sraplus. We paired data on catch histories, swept area ratio (SAR), and fisheries management index (FMI) with estimates of stock status from the RAM legacy stock assessment database. We then trained a model of the general form $log(status) \sim normal(variable,\sigma)$ for each of these three data types. Given values of these variables for a new fishery, sraplus uses the fitted model to generate posterior predictive distributions of stock status based on these data, which can then be used as priors on stock status by sraplus for new fisheries. For example, given data on SAR or FMI scores, together with a catch history, sraplus uses these regressions to convert those SAR and FMI values into priors on B/B~MSY~ or F/F~MSY~ in the most recent year of the fishery usable by sraplus (See Supplementary Information). All prior regression models where tested by out-of-sample predictive power, and where competing models were considered the final model was chosen by leave-on-out validation [@vehtari2017]. The final models are intended as a reasonably robust means of translating available data (catch histories, FMI, and SAR values) into a form usable by sraplus.

<!-- Given the scope of this analysis, we do not claim that the regressions used here are the best possible model relating these data with the fishery status indicators of interest. Rather, each regression was tested to ensure that it is unlikely that, given the same data, an alternative model would perform substantially better than those presented here. -->

@@

<!-- We first present a case study demonstrating how different kinds of data can lead to different conclusions about stock status. From there, we assess the performance of models fit using different kinds of broadly available data: combinations of catches, effort, Fisheries Management Index scores, and swept area ratio (SAR) values (Table.\@ref(tab:data)). -->

@@

We selected `r n_distinct(exs$stockid)` stocks for which we have stock specific FMI and SAR scores. We then paired effort data at the resolution of year, country, and FAO statistical area from Rousseau et al. 2019 [@rousseau2019] to each stock. As a benchmark, we first estimated stock status for these case study fisheries using the CMSY [@froese2017] method, as this has become one of the most widely used catch-only models currently available. We then used stock-specific data on SAR and FMI to generate priors on F/F~MSY~ for each of the stocks, which were then passed to sraplus. Lastly, we used the reconstructed effort data [@rousseau2019] to create an index of abundance for each stock, and estimated stock status by fitting to this index while using priors on fishing mortality rates informed by each stock's FMI and SAR values. While CMSY systemically overestimated fishing mortality rates and underestimated stock status, use of the SAR, FMI, and effort data produced substantially more accurate results for both B/B~MSY~ and F/F~MSY~ (Fig.\@ref(fig:cs-plot)).
